\documentclass[11pt]{article} %

\usepackage{amsmath}
\usepackage{comment}

\setlength{\oddsidemargin}{-0.15in}
\setlength{\topmargin}{-0.5in}
\setlength{\textwidth}{6.5in}
\setlength{\textheight}{9in}

\newcommand{\cP}{{\cal P}}
\newcommand{\IN}{{\bf N}}
\newcommand{\IZ}{{\bf Z}}
\newcommand{\IR}{{\bf R}}
\newcommand{\IQ}{{\bf Q}}
\newcommand{\points}[1]{{\it (#1 Points)}}
\newcommand{\qed}{{\hfill {\rm QED}}}

\begin{document}
%\maketitle




\noindent
{\Large Math 211 \qquad Homework 1} {\hfill \bf Name: Kelvin Abrokwa-Johnson}







\medskip\noindent
{\bf 1} (5 points) Find the point of intersection of the lines:
$x_1 + 2x_2 = -13, 3x_1 - 2x_2 = 1$.

Solution: To find the point of intersection, we solve the
system with augmented matrix
$$
\left[
\begin{matrix}
1 & 2 & -13 \cr
3 & -2  & 1
\end{matrix}
\right]
$$
First we row reduce:
$$
\left[
\begin{matrix}
1 & 2 & -13 \cr
3 & -2  & 1
\end{matrix}
\right]
$$
We subtract 3 times the first row from the second row to get:
$$
\left[
\begin{matrix}
1 & 2 & -13 \cr
0 & -8  & 40
\end{matrix}
\right]
$$
And then simplify:
$$
\left[\begin{matrix}
1 & 2 & -13 \cr
0 & 1  & -5 
\end{matrix}
\right]
$$
We can immediately see that  $x_2 = -5$. We plug that back into the first equation to solve for $x_1$:
\begin{align*}
x_1 &= -13 - 2x_2 \\
       &= -13 - 2(-5) \\
       &= -3
\end{align*}
And so our final solutions are:
\begin{align*}
x_1 &= -3 \\
x_2 &= -5
\end{align*}
\vskip .5in










\medskip\noindent
{\bf 2} (5 points)
Solve the linear system:
\begin{eqnarray*}
x_1 - 5x_2 + 4x_3 &=& -3 \cr
2x_1 - 7 x_2 + 3x_3 &=& -2 \cr
-2x_1+x_2+7x_3 &=& -1 \cr
\end{eqnarray*}
Solution. Consider augmented matrix
$$
\left[
\begin{matrix}
1  & -5  & 4 & -3 \cr
2  & -7  & 3 & -2 \cr
-2 &  1  & 7 & -1
\end{matrix}
\right].
$$
We can row reduce here starting with the third equation. We can start by subtracting 2 times the first row from the second row and adding 2 times the first row to the the third row:
$$
\left[
\begin{matrix}
1 & -5 & 4 & -3 \cr
0 & 3 & -5 & 4 \cr
0 & -9 & 15 & -7
\end{matrix}
\right] 
$$
Next, we add 3 times the second row to the third row:
$$
\left[
\begin{matrix}
1 & -5 & 4 & -3 \cr
0 & 3 & -5 & 4 \cr
0 & 0 & 0 & 5
\end{matrix}
\right] 
$$
We can now see that the system is inconsistent. There is no solution.

\vskip .5in









\medskip\noindent
{\bf 3} (5 points)
Solve the linear
system with augmented matrix
$$
\left[
\begin{matrix}
2 & 0 & -6& -8\cr
0 & 1 & ~ 2& ~ 3\cr
3 & 6 & -2 & -4\cr
\end{matrix}
\right].
$$
Solution: We can use row reduction. We start by subtracting 3 times the first column from the third column to get:
$$
\left[
\begin{matrix}
1 & 0 & -3 & -4 \cr
0 & 1 & 2& 3 \cr
0 & 6 & 7 & 8
\end{matrix}
\right]
$$
And for the second column we can subtract 6 times the second row from the third row to get:
$$
\left[
\begin{matrix}
1 & 0 & -3 & -4 \cr
0 & 1 & 2& 3 \cr
0 & 0 & -5 & -10
\end{matrix}
\right]
=
\left[
\begin{matrix}
1 & 0 & -3 & -4 \cr
0 & 1 & 2& 3 \cr
0 & 0 & 1 & 2
\end{matrix}
\right]
$$
We immediately see that $x_3 = 2$. We can then solve for $x_2$:
\begin{align*}
x_2 &= 3 - 2x_3 \\
       &= 3 - 2(2) \\
       &= -1
\end{align*}
And for $x_1$:
\begin{align*}
x_1 &= -4 + 3x_3 \\
       &= -4 + 3(2) \\
       &= 2
\end{align*}
Our solutions are:
\begin{align*}
x_1 &= 2 \\
x_2 &= -1 \\
x_3 &= 2
\end{align*}










\medskip\noindent
{\bf 4} (5 points) Find all the values $h$ so that the linear system with the following
augmented matrix is consistent:Applying row reduction to:
$$
\left[
\begin{matrix}
1 & h & -5 \cr
2&-8&6\cr
\end{matrix}
\right].
$$
Solution. To simplify our calculations a bit we can divide the second row by 2 and then flip the rows to get:
$$
\left[
\begin{matrix}
1 & -4 & 3 \cr
1 & h & 6
\end{matrix}
\right]
$$
Using row reduction, we get:
$$
\left[
\begin{matrix}
1 & -4 & 3 \cr
0 & h+4 & 3
\end{matrix}
\right]
$$
We can see that if $h = -4$ then we have 0 = a non-zero which would render the system inconsistent. So, the system is consistent exactly when $h \neq -4$.
\vskip .5in








\medskip\noindent
{\bf 5} (5 points) Row reduce the following augmented matrix.
Determine the leading ones and pivot columns.
$$
\left[
\begin{matrix}
1&2&4&5\cr
2&4&5&4\cr
4&5&4&3
\end{matrix}
\right].
$$
Solution: We begin our row reduction by subtracting 2 times the first frow from the second row and subtracting 4 times the first row from the third row to get:
$$
\left[
\begin{matrix}
1 & 2 & 4 & 5 \cr
0 & 0 & -4 & -6 \cr
0 & -3 & -12 & -17
\end{matrix}
\right].
$$
To simplify we can divide the second row by $-3$ and switch the second and third rows to get:
$$
\left[
\begin{matrix}
1 & 2 & 4 & 5 \cr
0 & -3 & -12 & -17 \cr
0 & 0 & 1 & 2
\end{matrix}
\right]. = 
\left[
\begin{matrix}
1 & 2 & 4 & 5 \cr
0 & 1 & 4  & \frac{17}{3} \cr
0 & 0 & 1 & 2
\end{matrix}
\right].
$$
With reference to the original matrix (before we swapped rows), the leading ones are at $(1,1)$, $(2,3)$, and $(3,2)$ and the pivot columns are 1, 2, and 3.
\vskip .5in










\medskip\noindent
{\bf 6} (5 points) Find the solutions of the system with augmented matrix:
$$
\left[
\begin{matrix}
1&-2&-1&4\cr
-2&4&-5&6\cr
\end{matrix}
\right]
$$
Solution: Using row reduction, we add 2 times the first column to the second column to get:
$$
\left[
\begin{matrix}
1 &- 2 & -1 & 4\cr
0 & 0 & -7 & 14 \cr
\end{matrix}
\right]
=
\left[
\begin{matrix}
1 &- 2 & -1 & 4\cr
0 & 0 & 1 & -2 \cr
\end{matrix}
\right].
$$
We can immediately see that $x_3 = -2$ and since there is no leading one in for the second column $x_2$ is a free variable (we'll call it $t$).
Solving for $x_1$ we get:
\begin{align*}
x_1 &= 4 + x_3 + 2x_2 \\
       &= 4 + (-2) + t \\
       &= 2 + t
\end{align*}
And so our final solution is:
\begin{align*}
x_1 &= 2 + t \\
x_2 &= t \\
x_3 &= -2
\end{align*}
\vskip .5in








\medskip\noindent
{\bf 7} (5 points)  Find the solutions of the system with augmented matrix:
$$
\left[
\begin{matrix}
1 & 0 & -9 & 0 & 4 \cr
0 & 1 & 3 & 0 & -1 \cr
0 & 0 & 0 & 1 & -7 \cr
0 & 0 & 0 & 0 &  0 \cr
\end{matrix}
\right].
$$
Solution: We can immediately see that $x_4 = -7$ and that $x_3$ is a free variable. We plug these in to solve for $x_2$:
\begin{align*}
x_2 = -1 + t
\end{align*}
and for $x_1$:
\begin{align*}
x_1 = 4 + 9t
\end{align*}
And our final solution:
\begin{align*}
x_1 &= 4 + 9t \\
x_2 &= -1 + t \\ 
x_3 &= t \\
x_4 &= -7
\end{align*}
\vskip .5in



\medskip\noindent
{\bf 8} (5 points) Determine the conditions on $(h,k)$ so that the following system
has no solution, one solution, and infinitely many solutions.
$$
x_1-3x_2 = 1, \quad 2x_1 + hx_2 = k.
$$
Solution. The equations produce the following augmented matrix:
$$
\left[
\begin{matrix}
1 & -3 & 1 \cr
2 & h & k
\end{matrix}
\right]
$$
To row reduce we subtract 2 times the first row from the second row to get:
$$
\left[
\begin{matrix}
1 & -3 & 1 \cr
0 & h+6 & k-2
\end{matrix}
\right]
$$
From this reduction we can see that if $h = -6$ and $k \neq 2$, the leading one would be the last entry, which would make the system inconsistent. If $h = -6$ and $k = 2$ then there are infinite solutions because the second row would be all zeroes. And finally, $h \neq -6$ is one solution.
\end{document}

